---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# viscomp

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/georgiosseitidis/viscomp.svg?branch=main)](https://app.travis-ci.com/github/georgiosseitidis/viscomp)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/georgiosseitidis/viscomp?branch=main&svg=true)](https://ci.appveyor.com/project/georgiosseitidis/viscomp)
[![Codecov test coverage](https://codecov.io/gh/georgiosseitidis/viscomp/branch/main/graph/badge.svg)](https://app.codecov.io/gh/georgiosseitidis/viscomp?branch=main)
<!-- badges: end -->

The goal of **viscomp** is to provide several visualization tools for exploring the behavior of the components in a network meta-analysis of complex interventions: 

 * components descriptive analysis
 * heat plot of the two-by-two component combinations 
 * leaving one component combination out scatter plot 
 * violin plot for specific component combinations' effects
 * density plot for components' effects  
 * waterfall plot for the interventions' effects that differ by a certain component combination 
 * network graph of components
 * rank heat plot of components for multiple outcomes.


## Installation

You can install the development version of **viscomp** like so:

``` r
install.packages("devtools")
devtools::install_github("georgiosseitidis/viscomp")
```

## How to use viscomp

We illustrate how to use **viscomp** in the vignette:

```{r}
vignette("viscomp", package = "viscomp")
```

## How to cite **viscomp**

```{r}
citation(package = "viscomp")
```
